\section{Design methodology and verification setup}
\label{sec:method}

\subsection{Sizing against the real load}
\label{sec:loading}

The decisive step of the design was recognizing what a delay stage actually
drives inside the grid. A $\ttau{2}$ tap feeds the latch inputs of an entire
row, a $\ttau{1}$ tap those of an entire column, and both feed the next delay
stage, whose input capacitance scales with the stage sizing itself. After
fan-out equalization the equivalent load is about twenty unit-inverter inputs
per tap.

We learned the cost of ignoring this directly. Stage sizing derived from a
standalone sweep with an idealized load predicted
$\tlsb=\SI{16.6}{\pico\second}$; the same cells assembled into a row with raw
tap loading measured \SI{28.6}{\pico\second}
(Section~\ref{sec:1dvs2d}). Nearly a factor of two in resolution, lost
entirely to loading. All final sizing was therefore done against a replica
load of ten latch inputs plus one downstream stage, swept in the slowest
corner (SS) so that the slow line stays reachable everywhere. This analysis
also moved the architecture itself: an earlier $k=4$ operating point
($\ttau{1}/\ttau{2}=60/\SI{45}{\pico\second}$) required stage delays below the
achievable window and was abandoned in favor of $k=6$.

\subsection{Per-corner calibration}
\label{sec:calmethod}

Process corners shift both line delays together, but not perfectly, so $\tlsb$
breathes with the corner. The MOSCAP banks of Section~\ref{sec:delaycell}
absorb this. The trim table was generated by an exhaustive probe: all
64 on/off combinations of the six bank gates were simulated per corner (320
short transients), measuring $\ttau{1}$ and $\ttau{2}$ directly on the tap
nets. The configuration with $\tlsb$ closest to \SI{15}{\pico\second} wins;
ties resolve toward $\ttau{1}$ closest to \SI{90}{\pico\second}. The winning
codes (Appendix~\ref{app:trim}) were then validated end-to-end with full
staircase sweeps. An independent hand derivation of the per-corner codes
agreed within one bank step in every corner. Calibration changes gate voltages
only, so the transistor inventory and $\sumw$ are identical in every corner.

\subsection{Testbench and measurement method}
\label{sec:testbench}

All system-level results come from the unmodified course testbench
\texttt{TB\_TDC} \cite{ee4615tb}: stimulus generation in \texttt{TDC\_in}
(RESET, START, STOP through $3\times$ minimum inverters), the DUT wrapper, and
\texttt{TDC\_out} loading every thermometer output with two $2\times$
inverters and \SI{1}{\femto\farad}. The DUT receives its supply through a
dedicated rail with a series ammeter, while the stimulus buffers run from a
separate testbench rail, so the measured energy contains only the converter.
Energy per conversion integrates the supply current over each \SI{5}{\nano\second}
conversion window, $E=\Vdd\int |i|\,dt$.

The characterization is automated in OCEAN. Per corner, the input delay is
swept from \SI{-15}{\pico\second} in \SI{1.5}{\pico\second} steps (one tenth of
the LSB) until code 31 is reached, recording every code transition; the sweep
then continues about 15\% further for the energy average. From the transition
table the scripts extract the LSB (mean step width), DNL, INL, offset, energy,
ENOB and the sign-off checks: no missing codes, monotonicity, and an intact
thermometer code at every point. The five corners run as parallel,
isolated Spectre jobs (1670 transients in the corner campaign). Two INL
conventions appear in the outputs: the course script reports the half-range
value $(\mathrm{INL}_{\max}-\mathrm{INL}_{\min})/2$, while the plots show the
endpoint-referenced per-code curve. We quote the course convention for
sign-off and plot the endpoint curve, stating both where they differ.
