\section{Conclusion}
\label{sec:conclusion}

We built and verified a 5-bit two-dimensional Vernier TDC in TSMC
\SI{180}{\nano\meter} BCD that meets every project requirement with margin.
With a single fixed trim setting the converter resolves
\SIrange{13.1}{18.2}{\pico\second} across all five process corners with no
missing codes (worst DNL \num{0.90} LSB); per-corner MOSCAP trimming
re-centers the LSB to \SIrange{15.0}{15.85}{\pico\second}, a six times tighter
spread, without changing a single device. The core consumes
\SI{11.7}{\pico\joule} per conversion in the typical corner at a figure of
merit of \SI{0.50}{\pico\joule} per step, within a total transistor width of
\SI{1305.6}{\micro\meter}. The arbiter contributes an offset of 1\% of an LSB
and a dead zone bounded below \SI{1}{\femto\second}, so the resolution is set
by the architecture, not by the decision element.

The evidence supports three conclusions beyond the headline numbers. First,
delay-cell sizing for Vernier grids must be done against the in-grid load;
the measured 1-D row shows the LSB doubling when tap loading is ignored.
Second, fan-out equalization with dummy latches is what makes the grid
thresholds usable; it costs 29 inactive cells and buys corner-independent
linearity. Third, the bijective $k=6$ mapping removes all combining logic,
and the only systematic linearity signature left, the period-6 wrap sawtooth,
is architectural and identical in every corner.

\paragraph{Recommendations.}
Three extensions would add the most value. A DLL-style background loop could
set the trim codes automatically, replacing the static calibration table. A
voltage-to-time front end would let the TDC quantize voltages directly,
addressing the course bonus. And a targeted capacitive correction at the four
column-wrap codes could suppress the period-6 DNL pattern that calibration
cannot reach. Layout and extracted re-simulation remain the natural next step
toward silicon; the schematic-level margins reported here (\SI{4}{\pico\second}
on resolution, four orders of magnitude on the dead zone) leave realistic room
for parasitic degradation.
